---
title: "Colocalization proportion of trans-eQTLs with GWAS loci"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')

# load packages -----
rm(list = ls())
library(tidyverse)
library(data.table)
```

## Goal

To look at what the proportion of trans-eQTLs that are colocalized with each of the traits analyzed.


## Proportion of trans-eQTLs coloc with traits

```{r fig.width = 10, fig.asp = 0.4, fig.align = "center"}
# I/O & paras -----
file_res_coloc_reg_prop1 <- '/project2/xuanyao/llw/coloc/ukbb_coloc_blood_traits/data/coloc_region_prop_merged.txt'
file_res_coloc_reg_prop2 <- '/project2/xuanyao/llw/coloc/immune_traits/pmid_all/coloc_region_prop_merged.txt'
file_res_coloc_reg_prop3 <- '/project2/xuanyao/llw/coloc/ukbb_coloc_more_traits/all_trait/data/coloc_region_prop_merged.txt'


# read files -----
res_coloc_reg_prop1 <- fread(file_res_coloc_reg_prop1, header = TRUE)
res_coloc_reg_prop2 <- fread(file_res_coloc_reg_prop2, header = TRUE)
res_coloc_reg_prop3 <- fread(file_res_coloc_reg_prop3, header = TRUE)

res_coloc_reg_prop2$trait <- res_coloc_reg_prop2$Phenocode
res_coloc_reg_prop <- rbind(res_coloc_reg_prop1, res_coloc_reg_prop2, res_coloc_reg_prop3)


df_plt1 <- res_coloc_reg_prop1 %>%
  arrange(desc(propPvalColocMerg)) %>%
  pivot_longer(cols = c(nRegionPvalMerg), names_to = "if_merge", values_to = "Num_reg") %>%
  pivot_longer(cols = c(nRegionPvalColocMerg), names_to = "if_merge_coloc", values_to = "Num_reg_coloc") %>%
  pivot_longer(cols = c(propPvalColocMerg), names_to = "if_merge_prop", values_to = "prop") %>%
  mutate(if_merge = replace(if_merge, if_merge == "nRegionPvalMerg", "(trans)"),
         if_merge_coloc = replace(if_merge_coloc, if_merge_coloc == "nRegionPvalColocMerg", "(coloc)"),
         trait_type = paste("Blood", if_merge),
         trait_type_coloc = paste("Blood", if_merge_coloc),
         trait_type_prop = paste("Blood", if_merge_prop))

df_plt2 <- res_coloc_reg_prop2 %>%
  arrange(desc(propPvalColocMerg)) %>%
  pivot_longer(cols = c(nRegionPvalMerg), names_to = "if_merge", values_to = "Num_reg") %>%
  pivot_longer(cols = c(nRegionPvalColocMerg), names_to = "if_merge_coloc", values_to = "Num_reg_coloc") %>%
  pivot_longer(cols = c(propPvalColocMerg), names_to = "if_merge_prop", values_to = "prop") %>%
  mutate(if_merge = replace(if_merge, if_merge == "nRegionPvalMerg", "(trans)"),
         if_merge_coloc = replace(if_merge_coloc, if_merge_coloc == "nRegionPvalColocMerg", "(coloc)"),
         trait_type = paste("Autoimmune", if_merge),
         trait_type_coloc = paste("Autoimmune", if_merge_coloc),
         trait_type_prop = paste("Autoimmune", if_merge_prop))

df_plt3 <- res_coloc_reg_prop3 %>%
  arrange(desc(propPvalColocMerg)) %>%
  pivot_longer(cols = c(nRegionPvalMerg), names_to = "if_merge", values_to = "Num_reg") %>%
  pivot_longer(cols = c(nRegionPvalColocMerg), names_to = "if_merge_coloc", values_to = "Num_reg_coloc") %>%
  pivot_longer(cols = c(propPvalColocMerg), names_to = "if_merge_prop", values_to = "prop") %>%
  mutate(if_merge = replace(if_merge, if_merge == "nRegionPvalMerg", "(trans)"),
         if_merge_coloc = replace(if_merge_coloc, if_merge_coloc == "nRegionPvalColocMerg", "(coloc)"),
         trait_type = paste("Other", if_merge),
         trait_type_coloc = paste("Other", if_merge_coloc),
         trait_type_prop = paste("Other", if_merge_prop))

df_plt <- rbind(df_plt1, df_plt2, df_plt3)


# add trait order
df_plt$trait <- fct_inorder(factor(df_plt$trait))


y_lim <- max(unique(df_plt$Num_reg))
y_nudge <- 2


base_plt <- ggplot(df_plt, aes(x = trait)) +
  geom_bar(aes(y = Num_reg, fill = trait_type), stat = "identity", position = position_dodge(width = 1)) +
  geom_bar(aes(y = Num_reg_coloc, fill = trait_type_coloc), stat = "identity", position = position_dodge(width = 1)) +
  geom_point(aes(y = prop*y_lim, color = trait_type_prop),
             #position = position_dodge(width = 1),
             show.legend = FALSE,
             shape = 18, size = 2,
             position = position_nudge(y = y_nudge)) +
  geom_line(aes(y = prop*y_lim, group = trait_type_prop, color = trait_type_prop),
            #position = position_dodge(width = 1),
            show.legend = FALSE,
            position = position_nudge(y = y_nudge)) +
  labs(x = NULL, y = "Number of regions", fill = "Region", color = NULL)

base_plt +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_y_continuous(limits = c(0, y_lim),
                     sec.axis = sec_axis(~./y_lim, name = "Coloc Proportion")) +
  #scale_fill_brewer(palette = "Paired", direction = -1)
  scale_fill_manual(
    breaks = c(
      "Blood (trans)", "Blood (coloc)",
      "Autoimmune (trans)", "Autoimmune (coloc)", 
      "Other (trans)", "Other (coloc)"
    ),
    values = c("Blood (trans)" = "#d9eaf3", "Autoimmune (trans)" = "#e2f3d4", "Other (trans)" = "#fde0e0",
               "Blood (coloc)" = "#186191", "Autoimmune (coloc)" = "#24731f", "Other (coloc)" = "#9e1213")
  ) +
  scale_color_manual(
    breaks = c("Blood propPvalColocMerg", "Autoimmune propPvalColocMerg", "Other propPvalColocMerg"),
    values = c("Blood propPvalColocMerg" = "blue",
               "Autoimmune propPvalColocMerg" = "#18e518",
               "Other propPvalColocMerg" = "#ff0000")
  ) +
  theme_classic() +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10, face = "bold"),
        #legend.background = element_rect(color = "black", linetype = "dashed"),
        legend.key.size= unit(0.5, "cm"),
        
        axis.line = element_line(colour="black"),
        plot.margin=unit(c(10,5,5,5),"mm"),
        
        #legend.margin = margin(-0.5,0,0,0, unit="cm"),
        
        axis.text.x = element_text(angle = 60, hjust=1, vjust = 1, size = 8, color = "black"),
        axis.text.y = element_text(colour = "black", size = 10),
        axis.title.y = element_text(size = 12),
        axis.title.y.right = element_text(angle = 90),
        axis.title.x = element_text(size = 12))

```

Dots show the proportion. Light shade bars show the number of trans-eQTL loci for coloc. Dark shade bars show the coloc regions with the corresponding traits on x-axis. Color for trait types.


## Look into the coloc regions for each module

Next, I further looked into what are those coloc regions for each trait exactly, in terms of what are the correpsonding module that coloc with the trait, and how many coloc regions for each (module, trait) pair.


```{r results='hold', collapse=FALSE}
# I/O & paras -----
file_list_resColoc <- c(
  list.files(
    "/project2/xuanyao/llw/coloc/ukbb_coloc_blood_traits",
    "^.*coloc_reg_w_merged.txt$",
    full.names = TRUE, recursive = TRUE
  ) %>% set_names("blood"),
  list.files(
    "/project2/xuanyao/llw/coloc/immune_traits",
    "^.*coloc_reg_w_merged.txt$",
    full.names = TRUE, recursive = TRUE
  ) %>% set_names("immune"),
  list.files(
    "/project2/xuanyao/llw/coloc/ukbb_coloc_more_traits",
    "^.*coloc_reg_w_merged.txt$",
    full.names = TRUE, recursive = TRUE
  ) %>% set_names("other")
)
file_trait_type <- '/project2/xuanyao/llw/coloc/ukbb_coloc_blood_traits/ukbb_blood_traits.csv'

# read files -----
resColoc_all <- lapply(file_list_resColoc, fread, header = TRUE, drop = "Phenocode")
resColoc_all <- bind_rows(resColoc_all[(lapply(resColoc_all, nrow) %>% unlist()) > 0],
                          .id = "trait_type")
trait_type_blood <- fread(file_trait_type, header = TRUE, sep = ",")


# Figure data prep -----
## assign more specific blood traits groups -----
resColoc_all[resColoc_all$trait_type == "blood", "trait_type"] <- resColoc_all %>%
  filter(trait_type == "blood") %>%
  left_join(trait_type_blood, by = c("trait" = "Trait Abbreviation")) %>%
  pull(`GWAS Group`)

## summarize the number of un-merged and merged coloc regions for (module, trait) -----
resPlot <- resColoc_all %>%
  group_by(trait_type, trait, Module) %>%
  summarise(num_of_regions = n_distinct(Region),
            num_of_regions_merg = n_distinct(merged_region)) %>%
  ungroup()

## set module order by num of coloc regions and change labels -----
Module_order = resPlot %>%
  separate(Module, into = c(NA, "module_num"), sep = "module", remove = FALSE, convert = TRUE) %>%
  group_by(Module, module_num) %>%
  summarise(n_coloc_regions = sum(num_of_regions),
            n_coloc_regions_merg = sum(num_of_regions_merg),
            n_coloc_trait = n()) %>%
  ungroup() %>%
  arrange(desc(n_coloc_trait), desc(n_coloc_regions_merg), desc(n_coloc_regions))
resPlot$Module = factor(resPlot$Module,
                        levels = Module_order$Module,
                        labels = paste0(Module_order$module_num))

## set phenotype order within each trait type by num of coloc regions -----
pheno_order <- resPlot %>%
  group_by(trait_type, trait) %>%
  summarise(n_coloc_regions = sum(num_of_regions),
            n_coloc_regions_merg = sum(num_of_regions_merg),
            n_coloc_m = n()) %>%
  ungroup() %>%
  group_by(trait_type) %>%
  arrange(n_coloc_m, n_coloc_regions_merg, n_coloc_regions, .by_group = TRUE) %>%
  ungroup()
resPlot$trait = factor(resPlot$trait,
                           levels = pheno_order$trait)

## assign colocs to trait type -----
resPlot <- mutate(resPlot,
                  "trait_color" = recode(trait_type,
                                         "Platelets" = "#0028a1",
                                         "Red blood cells" = "#85192d",
                                         "White blood cells" = "#e89c31",
                                         "immune" = "#339900",
                                         "other" = "black"))
axis_text_color <- resPlot %>% slice(match(levels(trait), trait))


# tile plot of merged coloc regions for (module, trait) -----
base_plt <- ggplot(resPlot, aes(Module, trait)) +
  geom_tile(aes(fill = num_of_regions_merg)) +
  scale_y_discrete(labels = function(x) str_wrap(x, width = 10)) +
  scale_fill_gradientn(colors = RColorBrewer::brewer.pal(8, "Blues")[3:8],
                       na.value = "red",
                       limits = c(0, max(c(resPlot$num_of_regions, resPlot$num_of_regions_merg)))) +
  labs(y = NULL, x = "Module", fill = "Number of \n Coloc Regions")

base_plt +
  theme_classic() +
  theme(
    legend.title = element_text(size = 10, face = "bold"),
    legend.background = element_rect(color = "black", linetype = "dashed"),
    axis.text.y = element_text(colour = axis_text_color$trait_color, size = 8),
    axis.text.x = element_text(colour = "black", size = 8,
                               angle = 60, hjust = 1, vjust = 1)
  )
```



## Session info

```{r}
sessionInfo()
```

