---
title: "Run simulations under alternatives"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')

library(tidyverse)
```

## Goal

To simulate alternative z-scores and calculate p-values under various scenarios using three association tests: trans-pco, minP, PC1.


## Simulations of alternatives under various scenarios

- Scenarios under various sample size: 200, 400, 600, 800.

```{bash eval=FALSE}
Rscript --no-restore --no-save simulation/2_1_sim_alt.R \
N \
0.1 \
simulation/script_lambda0.1/ \
/project2/xuanyao/llw/simulation_lambda0.1/new_Sigma/Sigma-new_DGN_module29_K101.rds \
simulation/result/simulation.alt.N.lambda0.1.varb1e-3.K101.rds \
0.001 \
0.3 \
500 \
10000 \
1000
```


- Scenarios under various proportions of target genes in a module: 1%, 5%, 10%, 30%, 50%
```{bash eval=FALSE}
Rscript --no-restore --no-save simulation/2_1_sim_alt.R \
caus \
0.1 \
simulation/script_lambda0.1/ \
/project2/xuanyao/llw/simulation_lambda0.1/new_Sigma/Sigma-new_DGN_module29_K101.rds \
simulation/result/simulation.alt.caus.lambda0.1.varb1e-3.K101.rds \
0.001 \
0.3 \
500 \
10000 \
1000
```


- Scenarios under various genetic variances: 0.002, 0.003, 0.004, 0.005, 0.006

```{bash eval=FALSE}
Rscript --no-restore --no-save simulation/2_1_sim_alt.R \
var \
0.1 \
simulation/script_lambda0.1/ \
/project2/xuanyao/llw/simulation_lambda0.1/new_Sigma/Sigma-new_DGN_module29_K101.rds \
simulation/result/simulation.alt.var.lambda0.1.varb1e-3.K101.rds \
0.001 \
0.3 \
500 \
10000 \
1000
```


## Calculate power

Calculate power using empirical FDR correction for three tests, under various sample sizes for example.

```{bash eval=FALSE}
Rscript --no-restore --no-save simulation/3_cal_power.R \
simulation/result/simulation.alt.N.lambda0.1.varb1e-3.K101.rds \
simulation/result/simulation.null.lambda0.1.K101.rds \
simulation/result/power.N.lambda0.1.varb1e-3.K101.rds \
0.1 \
True
```


## Visualize power

Compare the power of three association tests: trans-pco, minP, PC1, under various scenarios.


```{r power plot}
plt_power <- function(file_power, xlab_name){
  res.alt = map_dfr(
  readRDS(file_power),
  ~as_tibble(.x, rownames = "model") %>%
    pivot_longer(!model, names_to = NULL, values_to = "power") %>%
    mutate("model" = str_extract(model, "\\d+.*\\d+$")),
  .id = "method"
)
res.alt$method = factor(res.alt$method, c("PCO", "PC1", "minp"), c("Trans-PCO", "PC1", "MinP"))


# point range & line plot ----
base_plt = ggplot(data = res.alt, aes(x = model, y = power, color = method, group = method)) +
  stat_summary(
    geom = "pointrange",
    fun.data = "mean_cl_normal",
    fun.args = list(conf.int = .95),
    position = position_dodge(width = 0.3),
    fatten = 1, size = 0.5
  ) +
  stat_summary(
    geom = "line",
    fun = "mean",
    position = position_dodge(width = 0.3),
    alpha = 0.5, show.legend = FALSE
  ) +
  labs(x = xlab_name, y = "Power", color = "Method")

base_plt +
  scale_colour_manual(
    breaks = c("Trans-PCO", "PC1", "MinP"),
    values = c("Trans-PCO" = "#85192d", "PC1" = "#1d349a", "MinP" = "#e89c31"),
    guide = guide_legend(override.aes = list(size = 0.3))
  ) +
  theme_classic() +
  theme(
    panel.grid.major.y = element_line(linetype = "dashed", size = 0.8),
    
    legend.position = "right",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 12, face = "bold"),
    #legend.background = element_rect(color = "black", linetype = "dashed"),
    legend.key.size= unit(0.5, "cm"),
    
    axis.line = element_line(colour="black"),
    #axis.line.y = element_blank(),
    axis.ticks.y = element_blank(),
    
    axis.text=element_text(colour = "black", size=12),
    axis.title.y = element_text(angle=90,vjust =2, size=14),
    axis.title.x = element_text(vjust = -0.2, size=14),
    
    plot.margin=unit(c(10,5,5,5),"mm")
  ) +
  coord_cartesian(xlim = c(1.2, n_distinct(res.alt$model) - 0.2))
}
```


- Power plot under various sample sizes

```{r results='hold'}
plt_power(
  '/project2/xuanyao/llw/simulation_lambda0.1/new_Sigma/power.N.lambda0.1.varb1e-3.K101.rds',
  "Sample Size"
)
```


- Power plot under various proportions of target genes in a module

```{r results='hold'}
plt_power(
  '/project2/xuanyao/llw/simulation_lambda0.1/new_Sigma/power.caus.lambda0.1.varb1e-3.K101.rds',
  "Causal Proportion"
)
```


- Power plot under various genetic variances

```{r results='hold'}
plt_power(
  '/project2/xuanyao/llw/simulation_lambda0.1/new_Sigma/power.var.lambda0.1.varb1e-3.K101.rds',
  "Genetic Variance"
)
```


## Session info

```{r}
sessionInfo()
```


