---
title: "Run null simulations"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = '../')
```

## Goal

Run null simulations to show trans-pco and the other two methods, i.e. minP and PC1, are well-calibrated.


## Run three association tests on simulated null z-scores

First, load required packages.

```{r}
rm(list = ls())
library(mvtnorm)
source("/home/liliw1/Trans/plot/theme_my_pub.R")
```


Give input files and parameters. To make it simple (and fast), here I simulate only $10^3$ simulations (in the paper, $10^7$ tests were simulated).

```{r}
n_sim <- 10^3
```


```{r}
# I/O & paras -----
dir_pco <- 'simulation/script_lambda0.1/'
file_Sigma <- '/project2/xuanyao/llw/simulation_lambda0.1/new_Sigma/Sigma-new_DGN_module29_K101.rds'
```


Read files and source pco test scripts.

```{r}
# read files -----
# source pco test
source(paste0(dir_pco, "ModifiedPCOMerged.R"))
source(paste0(dir_pco, "liu.R"))
source(paste0(dir_pco, "liumod.R"))
source(paste0(dir_pco, "davies.R"))
dyn.load(paste0(dir_pco, "qfc.so"))
source(paste0(dir_pco, "ModifiedSigmaOEstimate.R"))

Sigma <- as.matrix(readRDS(file_Sigma))
K <- dim(Sigma)[1]
```


Simulate null z-scores using the given $\Sigma$ of a gene module from real dataset.

```{r}
# simulate null z-scores -----
set.seed(123)

p_null_all  <- list()
z_null <- rmvnorm(n_sim, rep(0, K), Sigma)
```


Run pco test and calculate p-values,

```{r}
# prepare eigenvalues and eigenvectors as input for three methods -----
SigmaO <- ModifiedSigmaOEstimate(Sigma)
eigen_res <- eigen(Sigma)
eigen_lamb <- eigen_res$values
eigen_vec <- eigen_res$vectors

# PCO
p_null_all$'p.null.PCO' <- ModifiedPCOMerged(
  Z.mat = z_null, Sigma = Sigma, SigmaO = SigmaO
) |> as.numeric()

cat("PCO done. \n\n")
```


Run pc1 test and calculate p-values,

```{r}
# PC1
PC1 <- z_null %*% eigen_vec[, 1]
p_null_all$'p.null.PC1' <- 2*pnorm(-abs(PC1/sqrt(eigen_lamb[1])))|> as.numeric()

cat("PC1 done. \n\n")
```


Run minp test and calculate p-values,

```{r}
# univariate minp
p_null_all$'p.null.minp' <- apply(z_null, 1, function(x) min(1-pchisq(x^2, 1))*length(x) )

cat("minp done. \n\n")
```

Finally, take a look at the calculated p-values on null z's.

```{r}
str(p_null_all)
```


## QQ-plot of null p-values by three association tests

First, load required packages.

```{r}
rm(list = ls())
library(tidyverse)
source("/home/liliw1/Trans/plot/theme_my_pub.R")
```

Files of simulated null p-values,

```{r}
# I/O & paras -----
file_dat_null <- "/project2/xuanyao/llw/simulation_lambda0.1/new_Sigma/simulation.null.lambda0.1.K101.rds"
ci_level <- 0.95
```


```{r}
# read files -----
p_null_all <- readRDS(file_dat_null)



# organize data -----
## reset obs p-values that are 0 to a fixed value, here I use the min non-zero p/10 -----
input <- as_tibble(p_null_all)
input[input == 0] <- min(input[input != 0])/10

## number of samples -----
n <- nrow(input)

## expected null p -----
expected <- seq(1, n) / (n+1)
lexp <- -log10(expected)

## order statistic of null p -----
ci_l <- -log10( qbeta(p = (1 - ci_level) / 2, shape1 = 1:n, shape2 = n:1) )
ci_r <- -log10( qbeta(p = (1 + ci_level) / 2, shape1 = 1:n, shape2 = n:1) )

## obs -----
observed <- apply(input, 2, sort) %>% as.data.frame()
lobs <- -log10(observed)


## take only a subset of null p's, to save image space -----
ind_sub <- c(
  1:sum(lexp > 4),
  seq(from = sum(lexp > 4), to = sum(lexp > 2), length.out = 2000) %>% ceiling(),
  seq(from = sum(lexp > 2), to = n, length.out = 2000) %>% ceiling()
)


# data for plt -----
df_plt <- cbind(data.frame(x = lexp, ci_l = ci_l, ci_r = ci_r), lobs) %>%
  slice(ind_sub) %>%
  pivot_longer(-c(x, ci_l, ci_r), names_to = "Type", values_to = "y")

## set group order in plt -----
group_order <- c("p.null.PCO", "p.null.PC1", "p.null.minp")
group_label <- c("Trans-PCO", "PC1", "MinP")

group_order <- if(is.null(group_order)) unique(df_plt$Type) else group_order
group_label <- if(is.null(group_label)) group_order else group_label
df_plt$Type <- factor(df_plt$Type, levels = group_order, labels = group_label)


# QQ-plot of null p-values by three association tests -----
base_plt <- ggplot(df_plt, aes(x = x, y = y, group = Type)) +
  geom_ribbon(aes(ymin = ci_l, ymax = ci_r), fill = "#e5e5e5", color = "#e5e5e5") +
  geom_abline(slope = 1, intercept = 0, color = "#595959", size = 0.7) +
  geom_point(aes(color = Type), size = 0.5) +
  labs(x = bquote(Expected -log[10]~italic((P))),
       y = bquote(Observed -log[10]~italic((P))),
       color = NULL)
base_plt +
  scale_color_manual(
    values = c("#85192d", "#0028a1", "#e89c31"),
    guide = guide_legend(override.aes = list(size = 2))
  ) +
  theme_my_pub() +
  theme(
    panel.grid.major.x = element_line(linetype = "dotted"),
    panel.grid.major.y = element_line(linetype = "dotted"),
    
    legend.background = element_blank(),
    legend.position = "right",
    
    axis.title = element_text(size = 14), 
    axis.text = element_text(colour = "black", size = 12)
  )
```



## Session info

```{r}
sessionInfo()
```


